Hazardous asteroids forecast via Markov random fields

### 2021-12-08 Presentation/Presentation.tex
PUB

\section{Introduction}

@@

\section{Preliminary analysis}

@@

\section{Probabilistic models}

@@

\section{Other ML algorithms}
\begin{frame}{Random Forest}
\begin{columns}
\column{0.5\textwidth}
  \begin{figure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{Pic/RF_confusion.pdf}
  \end{figure} 
  \column{0.5\textwidth}
  \begin{figure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{Pic/ROC_RF.pdf}
  \end{figure}
\end{columns}
\end{frame}

@@

\begin{frame}{forecast performances vs intepretability}
\begin{itemize}
\item The mgm algorithm is not the best one in term of performances, but it provides the connections between the features. On the other side, except for the variable importance in RF, the other are black box one
\item The mgm model, as the other graphical model is open to a true scientific validation, the other not. 
\item The probabilistic models lack in the forecast is definitely compensated by their interetability 
\item This is meaningful since this two features are in conflict 
\item The probabilistic models provide a good trade-off between intepretability and forecast performances, as long as one is interest to produce a really scientific result (e.g if the only aim is the forecast the RF is definitely better. However how long one can trust to the RF result ?)
\end{itemize}
\end{frame}

\begin{frame}[t,allowframebreaks]
\frametitle{References}
\printbibliography
\end{frame}

>>>>

\section{Intro}

@@

\section{Prelim. analysis}

@@

\section{Prob. models}

@@

\section{Other ML alg.}
\begin{frame}{Random Forest}
\begin{columns}
\column{0.5\textwidth}
  \begin{figure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{Pic/RF_confusion.pdf}
  \end{figure} 
  \column{0.5\textwidth}
  \begin{figure}[b]{\textwidth}
    \includegraphics[width=\textwidth]{Pic/ROC_RF.pdf}
  \end{figure}
\end{columns}
\end{frame}

@@

\begin{frame}{Conclusions: forecast performances vs intepretability}
\begin{itemize}
\item The mgm algorithm is not the best one in term of performances, but it provides the connections between the features. On the other side, except for the variable importance in RF, the other are black box one
\item The mgm model, as the other graphical model is open to a true scientific validation, the other not. 
\item The probabilistic models lack in the forecast is definitely compensated by their interetability 
\item This is meaningful since this two features are in conflict 
\item The probabilistic models provide a good trade-off between intepretability and forecast performances, as long as one is interest to produce a really scientific result (e.g if the only aim is the forecast the RF is definitely better. However how long one can trust to the RF result ?)
\end{itemize}
\end{frame}

\section{SI}

\begin{frame}
\begin{center}
\frametitle{$\phi$ coefficent}
\begin{table}[]
\begin{tabular}{c|c|c}
\multicolumn{1}{l|}{\textbf{}} & \textbf{Actual - N} & \textbf{Actual - P} \\ \hline
\textbf{Predicted - N}         & \#TP                & \#FN                \\ \hline
\textbf{Predicted - P}         & \#FP                & \#TP               
\end{tabular}
\end{table}
\begin{equation*}
\phi=\dfrac{TP\times TN - FP\times FN}{\sqrt{\left(TP+FP\right)\left(TP+FN\right)\left(TN+FP\right)\left(TN+FN\right)}}
\end{equation*}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Receiver operating characteristic}
\begin{columns}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=1\linewidth]{Pic/ROC/ROC_A.png}
\includegraphics[width=1\linewidth]{Pic/ROC/ROC_B.png}\\
\end{center}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_A_R.png}
\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_B_R.png}\\
\end{center}
\end{columns}
\begin{center}
Images taken from \cite{ROC_fig}
\end{center}
\end{frame}

\begin{frame}
\frametitle{Receiver operating characteristic}
\begin{columns}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=1\linewidth]{Pic/ROC/ROC_C.png}
\includegraphics[width=1\linewidth]{Pic/ROC/ROC_D.png}\\
\end{center}
\column{0.5\textwidth}
\begin{center}
\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_C_R.png}
\includegraphics[width=0.5\linewidth]{Pic/ROC/ROC_D_R.png}\\
\end{center}
\end{columns}
\begin{center}
Images taken from \cite{ROC_fig}
\end{center}
\end{frame}

\begin{frame}{Factor analysis of mixed data - FAMD}
\begin{center}
r(z,k) correlation coefficient (z and k quantitative) \\
\end{center}
\begin{center}
$\eta^{2}$(z,q) correlation ratio (z quantitative and q qualitative)
\end{center}
\begin{center}
PCA $\rightarrow$ $max\sum_{k}r^{2}(z,k)$
\end{center}
\begin{center}
MCA $\rightarrow$ $max\sum_{q}\eta^{2}(z,q)$
\end{center}
\begin{center}
FAMD $\rightarrow$ $max\sum_{k}r^{2}(z,k)+max\sum_{q}\eta^{2}(z,q)$
\end{center}
\end{frame}

\begin{frame}{mgm algorithm}
\begin{equation}
P(X_{s}|X_{\backslash s})=\exp\left\lbrace E_{s}(X_{\backslash s})\phi_{s}\left(X_{s}\right)+B_{s}(X_{s})-\Phi\left(X_{\backslash s}\right) \right\rbrace
\end{equation}
\begin{center}
$\phi_{s}$ function of sufficient statistics $B_{s}$ base measure
\end{center}

\begin{equation}
\begin{split}
P(X)= & \exp \left\lbrace \sum_{s\in V}\theta_{s}\phi_{s}(X_{s})+\sum_{s\in V}\sum_{r\in N(s)}\theta_{s,r}\phi_{s}(X_{s})\phi_{r}(X_{r})+...+\\&\sum_{r_{1},...,r_{k}\in C}\theta_{r1,...,rk}\prod_{j=1}^{k}\phi_{r}_{j}\left(X_{r}_{j}\right)+\sum_{s\in V}B_{s}(X_{S})-\Phi(\theta) \right\rbrace
\end{split}
\end{equation}

\begin{equation}
\hat{\theta}=arg \min_{theta}\left\lbrace -\mathcal{L}(\theta,X)+\lambda ||\theta ||_{1}  \right\rbrace\quad ||\theta ||_{1}=\sum_{j=1}^{J}|\theta_{j}|
\end{equation}

\begin{frame}{GLASSO}
\begin{equation}
L_{pen}(K,\hat{\mu})=log\det(K)-tr(K\;S)-\rho||K|| 
\end{equation}
\begin{center}
$K=\Sigma^{-1}$
\end{center}
\begin{center}
S: empirical covariance matrix
\end{center}
\end{frame}

\begin{frame}{Parameters tuning}
\begin{equation}
p\left(\textbf{y}|\boldsymbol{\theta}\right) = \dfrac{1}{Z(\theta)}\exp \left( \sum_{c}\boldsymbol{\theta}^{T}_{c}\phi_{c}\left(\textbf{y}\right)\right)
\end{equation}
\begin{equation}
\mathcal{L}\left(\boldsymbol{\theta}\right):= \frac{1}{N}\sum_{i}\log p\left(\textbf{y}_{i}|\boldsymbol{\theta}\right)=\frac{1}{N}\sum_{i}\left[\sum_{c} \boldsymbol{\theta}^{T}_{c}\phi_{c}(y_{i})-\log Z\left(\boldsymbol{\theta}\right)\right]
\end{equation}
\end{frame}

\begin{frame}{Parameters tuning}
\begin{equation}
\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\frac{1}{N}\sum_{i}\left[\phi_{c}(y_{i})-\frac{\partial}{\partial\boldsymbol{\theta}_{c}}\log Z(\boldsymbol{\theta})\right]
\end{equation}

\begin{equation}
\begin{split}
\frac{\partial \log Z(\boldsymbol{\theta})}{\partial\boldsymbol{\theta}}=\mathbb{E}\left[\phi_{c}(\textbf{y})\right|\theta]= \sum_{\textbf{y}}\phi_{c}(\textbf{y})p(\textbf{y}|\boldsymbol{\theta})
\end{split}
\end{equation}
\begin{equation}
\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\left[\frac{1}{N}\sum_{i}\phi_{c}(y_{i})\right]-\mathbb{E}\left[\phi_{c}(\textbf{y})\right]
\end{equation}
\begin{equation}
\frac{\partial\mathcal{L}}{\partial\boldsymbol{\theta}_{c}}=\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]-\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
\end{equation}
\begin{equation}
\mathbb{E}_{p_{emp}}\left[\phi_{c}(\textbf{y})\right]=\mathbb{E}_{p_{(\cdot|\boldsymbol{\theta})}}\left[\phi_{c}(\textbf{y})\right]
\end{equation}
\end{frame}

\begin{frame}{mmod}
\begin{equation}
\begin{split}
f(i,y)=& p(i)(2\pi)^{-q/2}det(\Sigma)^{-1/2} \\
& exp\left[-\dfrac{1}{2}\left(y-\mu(i)\right)^{T}\Sigma^{-1}\left(y-\mu(i)\right)\right]
\end{split}
\label{gaussMix}
\end{equation}
\begin{equation}
\begin{split}
f(i,y) & = \exp\left\lbrace g(i)+\sum_{u}h^{u}(i)y_{u}-\dfrac{1}{2}\sum_{uv} y_{u}y_{v}k_{uv}\right\rbrace \\
&= \exp\left\lbrace g(i)+h(i)^{T}y-\dfrac{1}{2}y^{T}Ky \right\rbrace
\end{split}
\end{equation}

\begin{center}
where $g(i)$, $h(i)$ and $K$ are the canonical parameters
\end{center}

\end{frame}
\begin{frame}{mmod}
\begin{equation}
\begin{split}
K=&\Sigma^{-1} \\
h(i)=&\Sigma^{-1}\mu(i) \\
g(i)=&\log p(i) -\frac{1}{2}\log det (\Sigma) \\
&-\dfrac{1}{2}\mu(i)^{T}\Sigma^{-1}\mu(i)-\dfrac{q}{2}\log 2\pi
\end{split}
\end{equation}
\end{frame}

\begin{frame}{Graphical models}
\begin{equation}
p(x_{1},x_{2},...,x_{n})
\end{equation}
\begin{equation}
p(x_{1:V})=p(x_{1})p(x_{2}|x_{1})p(x_{3}|x_{2},x_{1})...p(x_{V}|x_{1:V-1})
\end{equation}
\begin{equation}
X  \perp Y| Z \iff  p(X,Y|Z) = p(X|Z)p(Y|Z)
\end{equation}
\begin{equation}
p(\textbf{x}_{1:V})=p(x_{1})\prod^{V}_{t=1}p(x_{t}|x_{t-1})
\end{equation}
\end{frame}

\begin{frame}{Graphical models}
\begin{theorem}[Hammersley-Clifford]
A positive distribution p(\textbf{y})>0 satisfies the CI properties of an indirect graph G iif p can be represented as a product of factor, one per maximal clique,  i.e.
\begin{equation}
p(\textbf{y}|\theta)= \dfrac{1}{Z(\theta)}\prod_{c \in C }\psi_{c}(\textbf{y}_{c}|\theta_{c})
\end{equation}
where C is the set of all the (maximal) cliques of G,  and Z($\theta$) is the partition function given by 
\begin{equation}
Z(\theta):= \sum_{y}\prod_{c\in C}\psi_{c}(\textbf{y}_{c}|\theta_{c})
\end{equation}
Note that this partition function is what ensures the overall distribution sums to 1
\end{theorem}
\end{frame}
\begin{frame}{Graphical models}
\begin{equation}
p(y|\theta)=\dfrac{1}{Z(\theta)} exp\left(-\sum_{c}E(y_{c}|\theta_{c})\right)
\end{equation}
\begin{equation}
\psi_{c}(y_{c}|\theta_{c})=exp\left(-E(y_{c}|\theta_{c})\right)
\end{equation}
\end{frame}

\begin{frame}{Information theory}
\begin{equation}
H(X)=-\sum_{x \in X} p(x)log p(x)
\end{equation}
\begin{equation}
H(X,Y)=-\sum_{x\in \chi}\sum_{y\in \mathcal{Y}}p(x,y)\log p(x,y)
\end{equation}
\begin{equation}
\begin{split}
H(X|Y)=& \sum_{x\in \mathcal{X} }p(x)H(Y|X=x)\\
=& -\sum_{x\in \mathcal{X}}p(x)\sum_{y\in \mathcal{Y}}p(x,y)\log p(y|x) \\
=& -E\log p(Y|X)
\end{split}
\end{equation}
\begin{equation}
H(X,Y)=H(X)+H(Y|X)
\end{equation}
\begin{equation}
D(p||q)=\sum p(x)\log\frac{p(x)}{q(x)} \quad D(p||q)\geq 0
\end{equation}

\begin{frame}{Information theory}
\begin{equation}
\begin{split}
I(X;Y)&=\sum\sump(x,y)\log\dfrac{p(x,y)}{p(x)p(y)}=D(p(x,y)||p(x)p(y)) \\
&=H(X)-H(X|Y)=H(Y)-H(Y|X)
\end{split}
\end{equation}
\begin{equation}
I(X_{1},X_{2},...,X_{n};Y)=\sum^{n}_{i=1}I(X_{i};Y|X_{i-1},X_{i-2},...,X_{1})
\end{equation}
\begin{equation}
g(\alpha,\textbf{C},\textbf{S},f_{i})=MI(f_{i};\textbf{C})-\sum_{f_{s}\in S}\alpha(f_{i},f_{s},\textbf{C},\textbf{S})MI(f_{i};f_{s})
\end{equation}
\end{frame}

\begin{frame}{Ref}
\printbibliography
\end{frame}
